---
title: "xkcd Color Survey"
format: html
author: Nelly Tkachman
date: 2025-07-08
execute: 
  code-fold: true
  
  
---

TidyTuesday work for 7/8/25

The xKCD color survey asked people to name the colors that popped up on their screens.

**Acquire data**

```{r}
#| message: false
#| warning: false
#| code-fold: true

library(tidytuesdayR)
library(tidyverse)
library(plotly)


tuesdata <- tidytuesdayR::tt_load('2025-07-08')
answers <- tuesdata$answers
color_ranks <- tuesdata$color_ranks
users <- tuesdata$users

```
**Reds ranked by red value in RGB**
```{r}
#| message: false
#| warning: false
#| code-fold: true
rgb_matrix <- col2rgb(color_ranks$hex)
rgb_df <- as.data.frame(t(rgb_matrix))
red <- rgb_df %>% 
  mutate(rank = row_number()) %>% 
  right_join(color_ranks, join_by(rank)) %>% 
  filter(str_detect(color, "red$")) %>% 
  arrange(red, green, blue) %>% 
  mutate(row = row_number(), graphcolor = "red")

red %>% 
  ggplot(aes(x = row, y = graphcolor, label = color, fill = hex)) +
  geom_tile()+
  scale_fill_identity()+
  geom_text(color = "black")+
  coord_flip()


```
**Colors**
Based on the 2010 xkcd Color Name Survey, this graph shows colors that participants named using a descriptive version of the 11 basic color terms in English \nFor example 'dark red' is a red, 'reddish orange' is orange and 'yellowgreen' is green
```{r}
#| message: false
#| warning: false
#| code-fold: true
rgb_matrix <- col2rgb(color_ranks$hex)
rgb_df <- as.data.frame(t(rgb_matrix)) %>% mutate(rank = row_number())
colors_named <- color_ranks %>% 
  left_join(rgb_df, join_by(rank)) %>% 
  mutate(rainbow = case_when(
    str_detect(color,"red$") ~ "red", 
    str_detect(color,"orange$") ~ "orange",
    str_detect(color,"yellow$") ~ "yellow",
    str_detect(color,"green$") ~ "green",
    str_detect(color,"blue$") ~ "blue",
    str_detect(color,"purple$") ~ "purple",
    str_detect(color,"pink$") ~ "pink",
    str_detect(color,"brown$") ~ "brown", 
    str_detect(color,"black$") ~ "black",
    str_detect(color,"grey$") ~ "grey",
    str_detect(color,"black$") ~ "black",
    str_detect(color,"white$") ~ "white",
    TRUE ~ "other"
  )) %>% 
  mutate(category = case_when(rainbow == "red" ~ 1,
                              rainbow == "orange" ~ 2, 
                              rainbow == "yellow" ~ 3, 
                              rainbow == "green" ~ 4, 
                              rainbow == "blue" ~ 5, 
                              rainbow == "purple" ~ 6, 
                              rainbow == "pink" ~ 7, 
                              rainbow == "brown" ~ 8, 
                              rainbow == "grey" ~ 9, 
                              rainbow == "black" ~ 10,
                              rainbow == "white" ~ 11)) %>% 
  arrange(red, green, blue) %>% 
  group_by(rainbow) %>% 
  mutate(row = row_number())

p <- colors_named %>% 
  ggplot(aes(
    x = row, 
    y = reorder(rainbow, category), 
    text = paste0("<b>Color:</b> ", color, "<br>", "<bHex:</b> ", hex), 
    label = color, fill = hex))+
  geom_tile()+
  scale_fill_identity()+
  coord_flip()+
  labs(title = " Named colors using the 11 basic color terms", 
       subtitle = "Based on the 2010 xkcd Color Name Survey, this graph shows colors that participants named using a descriptive version of the 11 basic color terms in English \nFor example 'dark red' is a red, 'reddish orange' is orange and 'yellowgreen' is green,", 
       caption = "TidyTuesday July 8, 2025 \nData Source: xkcd Color Survey", 
       y = "Basic Color Term", 
       x = "")+
  theme(panel.background = element_rect(fill = "#fefdf8"), legend.position = "none")

ggplotly(p, tooltip = "text")
```
