---
title: "Biritsh Library Funding"
format: html
author: Nelly Tkachman
date: 2025-07-15
execute: 
  code-fold: true
editor: 
  markdown: 
    wrap: 72
---

British Library Funding from 1998 to 2023 Looking at the funding and the
funding adjusted for inflation

**Acquire date**

```{r}
#| message: false
#| warning: false
#| code-fold: true

library(tidytuesdayR)
library(tidyverse)

tuesdata <- tidytuesdayR::tt_load('2025-07-15')
bl_funding <- tuesdata$bl_funding
```

**Simple look at funding overtime** Red is nominal funding Blue line is
funding adjusted in 2025 Green line is adjusted funding that was done in
2016, based on an older inflation adjustment calculator

```{r}
#| message: false
#| warning: false
#| code-fold: true
bl_funding %>% 
  ggplot() +
  geom_line(aes(x = year, y = nominal_gbp_millions), color = "red")+
  geom_line(aes(x = year, y = total_y2000_gbp_millions), color = "blue")+
  geom_line(aes(x = year, y = year_2000_gbp_millions), color = "green")


```

**How much of funding is from grant-in-aid**

```{r}
#| message: false
#| warning: false
#| code-fold: true

bl_funding %>% 
  ggplot()+
  geom_bar(aes(x = year, y = nominal_gbp_millions), stat = "identity", fill = "red")+
  geom_bar(aes(x = year, y = gia_gbp_millions), stat = "identity", fill = "blue")
```

**British Library funding break down**

-   **gia** - grant-in-aid funding (core funding from UK government)
-   **services** - funding from service delivery within the remit of
    being a charity
-   **voluntary** - funding covering all voluntary contributions and
    donations, including the valuation of donated collection items
-   **investment** - funding from returns on savings and investments
-   **other** - anything that doesn't fit above

```{r}
#| message: false
#| warning: false
#| code-fold: true

bl_funding_pivoted <- bl_funding %>% 
  pivot_longer(cols = 2:17, names_to = "value", values_to = "quantity") %>% 
  mutate(source = case_when(
    value == "nominal_gbp_millions" ~ "total_reported_funding", 
    value == "gia_gbp_millions" ~ "reported_funding", 
    value == "voluntary_gbp_millions" ~ "reported_funding", 
    value == "investment_gbp_millions" ~ "reported_funding", 
    value == "services_gbp_millions" ~ "reported_funding", 
    value == "other_gbp_millions" ~ "reported_funding", 
    value == "year_2000_gbp_millions" ~ "2016_adjusted_funding", 
    value == "total_y2000_gbp_millions" ~ "total_adjusted_funding",
    value == "gia_y2000_gbp_millions" ~ "adjusted_funding", 
    value == "voluntary_y2000_gbp_millions" ~ "adjusted_funding",
    value == "investment_y2000_gbp_millions" ~ "adjusted_funding",
    value == "services_y2000_gbp_millions" ~ "adjusted_funding", 
    value == "other_y2000_gbp_millions" ~ "adjusted_funding", 
    value == "inflation_adjustment" ~ "inflation_adjustment",
    value == "percentage_of_y2000_income" ~ "percentage_of_y2000_income",
    value == "gia_as_percent_of_peak_gia" ~ "gia_as_percent_of_peak_gia"
  ))

bl_funding_pivoted %>% 
  filter(source == "reported_funding") %>% 
  ggplot(aes(x = year, y = quantity, fill = fct_reorder(value, quantity)))+ 
  geom_area()+
  labs(title = "British Library funding streams 1998 - 2023 (not inflation adjusted)")+
  ylab("GBP in millions")

bl_funding_pivoted %>% 
  filter(source == "adjusted_funding") %>% 
  ggplot(aes(x = year, y = quantity, fill = fct_reorder(value, quantity)))+ 
  geom_area()+
  labs(title = "British Library funding streams 1998 - 2023 (inflation adjusted)")+
  ylab("GBP in y2000 millions")

```
